---
title: "GOT"
author: "Sascha Berg"
date: "12/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(here)

# For text mining:
library(pdftools)
library(tidytext)
library(textdata) 
library(ggwordcloud)

# Note - Before lab:
# Attach tidytext and textdata packages
# Run: get_sentiments(lexicon = "nrc")
# Should be prompted to install lexicon - choose yes!
# Run: get_sentiments(lexicon = "afinn")
# Should be prompted to install lexicon - choose yes!

```

### Get the GOT-file
```{r get-document}
got_path <- here("data","got.pdf")
got_text <- pdf_text(got_path)
```


Getting a single page
```{r single-page}
got_p9 <- got_text[9]
got_p9
```
```{r split-lines}
got_df <- data.frame(got_text) %>% 
  mutate(text_full = str_split(got_text, pattern = '\\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) 
```


### Get the tokens (individual words) in tidy format

```{r tokenize}
got_tokens <- got_df %>% 
  unnest_tokens(word, text_full)

# See how this differs from `ipcc_df`
# Each word has its own row!
```

Counting the words
```{r count-words}
got_wc <- got_tokens %>% 
  count(word) %>% 
  arrange(-n)
got_wc
```

### Remove stop words:


```{r stopwords}
got_stop <- got_tokens %>% 
  anti_join(stop_words) %>% 
  select(-got_text)
```

Now check the counts again: 
```{r count-words2}
got_swc <- got_stop %>% 
  count(word) %>% 
  arrange(-n)
```


Getting rid of all the numbers 
```{r skip-numbers}
got_no_numeric <- got_stop %>% 
  filter(is.na(as.numeric(word)))
```


### A word cloud of GOT report words (non-numeric)


```{r wordcloud-prep}
length(unique(got_no_numeric$word))

#Including the top 100 most frequent words
got_top100 <- got_no_numeric %>% 
  count(word) %>% 
  arrange(-n) %>% 
  head(100)
```

```{r wordcloud}
got_cloud <- ggplot(data = got_top100, aes(label = word)) +
  geom_text_wordcloud() +
  theme_minimal()

got_cloud
```

Customizing:
```{r wordcloud-pro}
ggplot(data = got_top100, aes(label = word, size = n)) +
  geom_text_wordcloud_area(aes(color = n), shape = "diamond") +
  scale_size_area(max_size = 12) +
  scale_color_gradientn(colors = c("darkgreen","blue","red")) +
  theme_minimal()
```

"afinn": Words ranked from -5 (very negative) to +5 (very positive)
```{r afinn}
get_sentiments(lexicon = "afinn")
# Note: may be prompted to download (yes)

# Let's look at the pretty positive words:
afinn_pos <- get_sentiments("afinn") %>% 
  filter(value %in% c(3,4,5))

# Do not look at negative words in class. 
afinn_pos
```

bing: binary, "positive" or "negative"
```{r bing}
get_sentiments(lexicon = "bing")
```


```{r nrc}
get_sentiments(lexicon = "nrc")
```


### Sentiment analysis with afinn: 

First, bind words in `got_stop` to `afinn` lexicon:
```{r bind-afinn}
got_afinn <- got_stop %>% 
  inner_join(get_sentiments("afinn"))
```
Finding counts by sentiment ranking:
```{r count-afinn}
got_afinn_hist <- got_afinn %>% 
  count(value)

# Plot them: 
ggplot(data = got_afinn_hist, aes(x = value, y = n)) +
  geom_col()
```

Investigate some of the words in a bit more depth:
```{r afinn-2}
# What are these '2' words?
got_afinn2 <- got_afinn %>% 
  filter(value == 2)
```

```{r afinn-2-more}
# Check the unique 2-score words:
unique(got_afinn2$word)

# Count & plot them
got_afinn2_n <- got_afinn2 %>% 
  count(word, sort = TRUE) %>% 
  head(10) %>% 
  mutate(word = fct_reorder(factor(word), n))


ggplot(data = got_afinn2_n, aes(x = word, y = n)) +
  geom_col() +
  coord_flip()

```

Summarizing the sentiment for the report
```{r summarize-afinn}
got_summary <- got_afinn %>% 
  summarize(
    mean_score = mean(value),
    median_score = median(value)
  )
```

```{r}
head(got_summary)
```

As we can se the mean_score is -0.54, and therefore the sentiment analysis shows that the text is on the negative side.

### NRC lexicon for sentiment analysis

We can use the NRC lexicon to start "binning" text by the feelings they're typically associated with. As above, we'll use inner_join() to combine the GOT non-stopword text with the nrc lexicon: 

```{r bind-bing}
got_nrc <- got_stop %>% 
  inner_join(get_sentiments("nrc"))
```

Using `anti_join()`:

```{r check-exclusions}
got_exclude <- got_stop %>% 
  anti_join(get_sentiments("nrc"))

# View(ipcc_exclude)

# Count to find the most excluded:
got_exclude_n <- got_exclude %>% 
  count(word, sort = TRUE)

head(got_exclude_n)
```


Finding some counts 
```{r count-bing}
got_nrc_n <- got_nrc %>% 
  count(sentiment, sort = TRUE)

# And plot them:

ggplot(data = got_nrc_n, aes(x = sentiment, y = n)) +
  geom_col()
```

Counting by sentiment *and* word, then faceting:
```{r count-nrc}
got_nrc_n5 <- got_nrc %>% 
  count(word,sentiment, sort = TRUE) %>% 
  group_by(sentiment) %>% 
  top_n(5) %>% 
  ungroup()

got_nrc_gg <- ggplot(data = got_nrc_n5, aes(x = reorder(word,n), y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, ncol = 2, scales = "free") +
  coord_flip() +
  theme_minimal() +
  labs(x = "Word", y = "count")

# Show it
got_nrc_gg

# Save it
ggsave(plot = got_nrc_gg, 
       here("figures","ipcc_nrc_sentiment.png"), 
       height = 8, 
       width = 5)

```

```{r nrc-confidence}
conf <- get_sentiments(lexicon = "nrc") %>% 
  filter(word == "honor")

# Yep, check it out:
conf
```





